---
title: ''
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
  pdf_document:
    toc: true
    toc_depth: '3'
---
### Name: Richard Lam
### EID: rwl734
### GitHub Repo: https://github.com/RichardLam27/SDS315.git

\
\
\

```{r global_options, include = FALSE}

#importing libraries, datasets, and global code options

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE, 
                      warning = FALSE, 
                      message = FALSE
                      )
library(MatchIt)
library(tidyverse)
library(lubridate)
library(knitr)
library(rvest)
library(mosaic)

  
```

## Problem 1 - Armfold

### Part A

```{r}
armfold <- read.csv("armfold.csv")
armfold |>
   count(Sex)
armfold |>
  group_by(Sex) |>
  summarise(prop_left_on_top = mean(LonR_fold))
```
\

### Part B

```{r}
diff <- .4716981 - .4234234
cat("The observed difference is", diff, "\n")

```

\

### Part C
```{r}
# Calculate the sample proportions for males and females
male_prop <- mean(armfold$LonR_fold[armfold$Sex == "Male"])
female_prop <- mean(armfold$LonR_fold[armfold$Sex == "Female"])

n1 <- sum(armfold$Sex == "Male")
n2 <- sum(armfold$Sex == "Female")

SE <- sqrt((male_prop * (1 - male_prop)) / n1 + (female_prop * (1 - female_prop)) / n2)

z_star <- 1.96


ME <- z_star * SE


diff_prop <- male_prop - female_prop


CI_lower <- diff_prop - ME
CI_upper <- diff_prop + ME

cat("The 95% confidence interval for the difference in proportions is: [", CI_lower, ",", CI_upper, "]\n")


prop_test_result <- prop.test(c(sum(armfold$LonR_fold[armfold$Sex == "Male"]), 
                                  sum(armfold$LonR_fold[armfold$Sex == "Female"])), 
                                c(n1, n2), 
                                conf.level = 0.95, 
                                alternative = "two.sided")


cat("Result from prop.test():\n")
print(prop_test_result)

```
The standard error (SE) for the difference in proportions is: SE = sqrt((p1 * (1 - p1)) / n1 + (p2 * (1 - p2)) / n2), where p1 is the proportion of males who folded their left arm on top (0.4717), p2 is the proportion of females who folded their left arm on top (0.4234), n1 is the number of males (153), and n2 is the number of females (147). For a 95% confidence interval, the(z* used is 1.96, which is the standard normal distribution for a two-sided test. 

\

### Part D

If we were to repeat this study many times with different random samples from the same population, then we would expect that about 95% of the resulting confidence intervals would contain the true difference in proportions between males and females who fold their left arm on top.

\

### Part E

 The standard error represents how much the difference in sample proportions (between males and females) might vary just by chance from sample to sample. 
 
 \
 
### Part F
 
 The sampling distribution is the distribution of the difference in sample proportions (male minus female) from many random samples. The part that varies is the sample difference; the part that stays fixed is the true population difference.
 
 \
 
### Part G
 
 The Central Limit Theorem justifies using a normal distribution. It says that if the sample sizes are large enough, the sampling distribution of the difference in proportions will be approximately normal, even if the original data isn’t.
 
 \
 
### Part H
 
 I would say that they are technically right since our interval includes 0, we cannot be fully certain that there is a difference.
 
 \
 
Yes, the confidence interval would be different each time because random samples vary. But if we repeated the experiment many times, about 95% of the confidence intervals would contain the true difference in population proportions.

## Question 2: Get out the vote

### Part A

```{r}
# Load dplyr if not already loaded
turnout = read.csv("turnout.csv")
# Proportion of those receiving a GOTV call who voted in 1998
prop_gotv_voted <- turnout |>
  filter(GOTV_call == 1) |>
  summarise(prop_voted = mean(voted1998 == 1))

# Proportion of those not receiving a GOTV call who voted in 1998
prop_no_gotv_voted <- turnout |>
  filter(GOTV_call == 0) |>
  summarise(prop_voted = mean(voted1998 == 1))

# Print the proportions
cat("Proportion of GOTV call recipients who voted in 1998:", prop_gotv_voted$prop_voted, "\n")
cat("Proportion of non-recipients who voted in 1998:", prop_no_gotv_voted$prop_voted, "\n")

# 95% confidence interval for the difference in proportions (GOTV - non-GOTV)
test <- prop.test(
  c(sum(turnout$voted1998[turnout$GOTV_call == 1]), sum(turnout$voted1998[turnout$GOTV_call == 0])),
  c(sum(turnout$GOTV_call == 1), sum(turnout$GOTV_call == 0)),
  conf.level = 0.95
)

cat("95% Confidence Interval for the difference in proportions: ", test$conf.int, "\n")

```

\

### Part B

```{r}


summary_stats <- turnout |>
  group_by(voted1998) |>
  summarise(
    mean_voted1996 = mean(voted1996),
    mean_age = mean(AGE),
    mean_majorpty = mean(MAJORPTY)
  )

print(summary_stats)

summary_stats_gotv <- turnout |>
  group_by(GOTV_call) |>
  summarise(
    mean_voted1996 = mean(voted1996),
    mean_age = mean(AGE),
    mean_majorpty = mean(MAJORPTY)
  )

print(summary_stats_gotv)

age_gotv_test <- t.test(AGE ~ GOTV_call, data = turnout)
print(age_gotv_test)

majorpty_gotv_test <- t.test(MAJORPTY ~ GOTV_call, data = turnout)
print(majorpty_gotv_test)

voted1996_gotv_test <- t.test(voted1996 ~ GOTV_call, data = turnout)
print(voted1996_gotv_test)

```
In conclusion, I used t.test to compare the means of these variables between the two groups to check if these variables are different between the groups in order to see if they are confounding variables, which it turns out MAJORPTY, voted1996, and age are significantly different across the GOTV call groups, which means that they could be potential confounders.

\

### Part C

```{r}

match_model <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data = turnout, method = "nearest", ratio = 5)

matched_data <- match.data(match_model)

summary(match_model)

prop_gotv <- mean(matched_data$voted1998[matched_data$GOTV_call == 1])  
prop_no_gotv <- mean(matched_data$voted1998[matched_data$GOTV_call == 0])  

diff_proportions <- prop_gotv - prop_no_gotv

prop_test_result <- prop.test(c(sum(matched_data$voted1998[matched_data$GOTV_call == 1]), sum(matched_data$voted1998[matched_data$GOTV_call == 0])), 
                              c(sum(matched_data$GOTV_call == 1), sum(matched_data$GOTV_call == 0)))

cat("Proportion for GOTV call group: ", prop_gotv, "\n")
cat("Proportion for no GOTV call group: ", prop_no_gotv, "\n")
cat("Difference in proportions: ", diff_proportions, "\n")
cat("95% Confidence Interval for the difference in proportions: ", prop_test_result$conf.int, "\n")

```
After adjusting for key confounding variables such as prior voting behavior (voted1996), age, and party registration status (MAJORPTY) through the matching, we find evidence that receiving the GOTV call had a good effect on the likelihood of voting in the 1998 congress election. In the matched sample, individuals who received a GOTV call were about 7.85 percentage points more likely to vote compared to those who did not receive a call. This difference is statistically significant, with a 95% confidence interval ranging from approximately 1.05 to 14.66 percentage points. Because the covariates were well-balanced after matching, we can say that the matching worked and that the GOTV call worked and the campagn was effective.
